### 2016-01-28 paper/chemokine_supplement_revised.tex
In process of drafting Response 3

\documentclass[preprint,10pt,authoryear]{elsarticle}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

@@

@@

An overview of the model implementation is shown in Fig. 2.  The model is initialized with one virus-secreting cell placed in the middle of a 2-dimensional sheet of hexagonally-tiled epithelial cells (288,212 total cells).  The sheet measures 5$mm$ per side, which is large enough to contain the infection (Fig. 5).  The molecular grid (Text S1 1.3) is initialized with a resolution of 20$\mu m$.  Each time step of the model represents 6 seconds.

@@

\subsection{Modeling Decisions}

@@

The model simulates a chemokine gradient surrounding the infected focus (Fig. 5), based on the calculated per-infected cell secretion rate (Table 3) and known chemical parameters for a 10 kDa protein (Table 2).  T cell sensitivity depends on receptor density \citep{Desmetz2006} and this was assumed to be constant.

@@

% we hypothesize that it renders IP-10 ineffectual.  We hypothesize that this leads to the elevated RANTES secretion rates measured in aH5N1 compared to the other two strains (Table 3).  Because of this behavior IP-10 was not included in the aH5N1 model runs.

@@

\subsection{General sensitivity analysis}

@@

The model parameters were chosen from literature when available and estimated within plausible ranges otherwise (Tables 1 and 2).  Individual parameters were varied over ranges of plausible (and sometimes even implausible) values while the rest of the parameter set was held constant.  Each parameter was varied over all three influenza strains, creating three sets of sensitivity plots (Figures~\ref{fig:asensitivity}-\ref{fig:psensitivity}).

We then categorized the model parameters into one of three qualitative groups: parameters that do not affect the model's qualitative behavior (stable parameters), parameters that affect the model's quantitative peak infection size but do not affect final clearance, (peak change parameters), and those parameters that affect the final clearance of the infection (sensitive parameters) (Table 3).  Parameters are presented along with their range using the following format:  \\

@@

Stable parameters do not affect the outcome of the infection unless they are adjusted to values outside the realm of possibility.  Specifically, the plausible parameter values in this group are within a range that does not significantly affect the asymptotic behavior of the model.  We further split this group into two categories in Table 3, \textit{stable} parameters and \textit{bounded stable} parameters.  \textit{Bounded stable} parameters are only stable within a specific range of values, while \textit{stable} parameters seem to be stable over all the values that were tested.

@@

Compartmental modeling design suggests the removal of such stable parameters.  Due to the nature of the ABM, inclusion of these parameters is mechanistically necessary.  Thus, we include and parameterize them using the methods described in the main paper and Table 2.

@@

\textbf{Circulating T Cell Expected Age:} [ \textit{1h |  4d | 400d} ] - The expected age determines the decay rate of T cells in blood.

@@

\textbf{Chemotaxing T Cell Expected Age:} [ \textit{12m | 2h | 20h} ] - The expected age determines the decay rate of T cells in the epithelium.

@@

An explicit sensitivity analysis was not directly performed.  The range of the parameters are taken from the respective secretion rates of the aH5N1, sH1N1, and pH1N1 influenzas (Table 1).

@@

@@

\begin{figure}[!ht]
\begin{center}
 \includegraphics[width=\textwidth]{Figure_S-1}
 \end{center}
\caption{{\bf Varying T cell sensitivity to chemokine.}  H5N1 model results use RANTES  only, and sH1N1 and pH1N1 use both IP-10 and RANTES. Total number of incubating, secreting and apoptotic cells are plotted for each infection.  The sensitivity value specifies the minimum level of chemokine concentration required for T cells to detect it. } 
 \label{fig:sensitivity}
\end{figure}

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=\textwidth]{Figure_S2}
	\caption{\textbf{Effects of different chemokine combinations.}  A) aH5N1 does not stimulate an IP-10 response.  B-C) sH1N1 and pH1N1 show no significant difference between IP-10 alone versus IP-10 and RANTES combined.}
	\label{fig:chemokine}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=9in]{Figure_S3}
	\caption{\textbf{aH5N1 sensitivity analysis.} The sensitivity analysis results for 16 parameters applied to the aH5N1 model simulation.  In each subgraph, the thicker purple line shows the results of the model using the default parameter value.}
	\label{fig:asensitivity}
\end{center}
\end{figure}

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=9in]{Figure_S4}
	\caption{\textbf{sH1N1 sensitivity analysis.} The sensitivity analysis results for 16 parameters applied to the sH1N1 model simulation.  In each subgraph, the thicker purple line shows the results of the model using the default parameter value.  The divergence between Apoptosis Time values of 90m and 120m (upper-left plot) highlights an important threshold of infection control.}
	\label{fig:ssensitivity}
\end{center}
\end{figure}

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=9in]{Figure_S5}
	\caption{\textbf{pH1N1 sensitivity analysis.} The sensitivity analysis results for 16 parameters applied to the pH1N1 model simulation.  In each subgraph, the thicker purple line shows the results of the model using the default parameter value.}
	\label{fig:psensitivity}
\end{center}
\end{figure}

@@

%\begin{figure}[ht!]
%\begin{center}
%	\includegraphics[width=\textwidth]{Figure_S6}
%	\caption{\textbf{pH1N1 sensitivity analysis.} Model results and fits of the viral effectiveness rate, $E$.}
%	\label{fig:effectiveness}
%\end{center}
%\end{figure}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|  l  l  l  |}
  \hline
  Population & Description & Initial Value\\
  \hline
  $T$ & Healthy target epithelial cells & 1,000,000 \\
  $I_1$ & Virus-incubating cells & 0 \\
  $I_2$ & Virus-secreting cells & 0 \\
  $V$ & Free virus particles & 10,000 \\
  $F$ & Interferon quantity & 0 \\
  $C$ & Chemokine quantity (ng/mL) & 0 \\
  \hline
  \hline
  Parameter & Description & Value \\
  \hline
  $\beta$ & Viral infection rates &  \\
    & \hspace{2em} Avian & 5.3e-7 (PFU$\cdot$h)$^{-1}$ \\
    & \hspace{2em} Seasonal & 6.1e-7 (PFU$\cdot$h)$^{-1}$ \\
    & \hspace{2em} Pandemic &  2.7e-6 (PFU$\cdot$h)$^{-1}$\\
  $p$ & Viral production rates & \\
    & \hspace{2em} Avian & 0.20 (PFU/h) \\
    & \hspace{2em} Seasonal & 1.4 (PFU/h) \\
    & \hspace{2em} Pandemic & 18.3 (PFU/h) \\
  $e$ & Interferon strengths & \\
    & \hspace{2em} Avian & 1.0e-8 \\
    & \hspace{2em} Seasonal & 1.6e-6 \\
    & \hspace{2em} Pandemic & 3.4e-3 \\
  $\tau_2$ & Interferon production delays &  \\
    & \hspace{2em} Avian & 21.5 (h) \\
    & \hspace{2em} Seasonal & 23.6 (h) \\
    & \hspace{2em} Pandemic & 21.0 (h) \\
  $\delta$ & Virus-secreting cell decay rate & 0.6 (h$^{-1}$) \\
  $d$ & Chemokine decay rate & 1.386 (h$^{-1}$) \\
  $\tau_1$ & Viral incubation time & 10 (h) \\
  $\tau_3$ & Chemokine production delays & \\
    & \hspace{2em} IP-10 & 8 (h)\\
    & \hspace{2em} RANTES & 16 (h)\\
  \hline
\end{tabular}
\caption{\textbf{Parameters for Preliminary Model 1 (Equation 1)}.  All parameters and populations are taken from \citep{Mitchell2011} except for $C$, $d$, and $\tau_3$.  The value chosen for $d$ corresponds to a 30 minute half-life and values for $\tau_3$ were set from the observed chemokine data (Fig. 3).  Interferon ($F$) is an abstracted unitless quantity and thus $e$ is a unitless multiplier.}
\label{tab:dde}
\end{center}
\end{table}

\begin{table}[!ht]
\begin{center}
\begin{tabular}{ | c | c | c | }
  \hline                        
  Paramter & Value & Source \\
  \hline
  T(0) & $1e6$ & \citep{Mitchell2011} \\
  V(0) &  $1e4$ & \citep{Mitchell2011} \\
  $r_{cell}$ &  $5 \mu m$ & \citep{Miao2010a} \\
  $t_{rc}$ & $6s$ & \citep{Peters1983} \\
  $k_e$ & $6.4e-5/cell/day$ & \citep{Miao2010a} \\
  $\beta$ & $4.8e-7/cell/PFU$ & \citep{Mitchell2011} \\
  $p$ & $0.18 PFU/h$ & \citep{Mitchell2011} \\
  $\delta$ & $16.7 hours$ & \citep{Mitchell2011} \\
  $a$ & $67.7$ & ABM fits \\
  $b$ & $433.5$ & ABM fits \\
  \hline  
\end{tabular}
\caption{\textbf{Parameters for Preliminary Model 2 (Equation 2).}  Values are chosen to match the equations borrowed from both \citep{Mitchell2011} and \citep{Miao2010a}.  Values \textit{a} and \textit{b} were fit to multiple runs of a simplified CyCells ABM.}
\label{tab:supplement}
\end{center}
\end{table}

>>>>

%\documentclass[preprint,10pt,authoryear]{elsarticle}
\documentclass[preprint,10pt,authoryear]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
%\usepackage{amssymb}

@@

\usepackage{xr}
\externaldocument{chemokine_main_revised}

@@

An overview of the model implementation is shown in Fig.~\ref{fig:modelchart}.  The model is initialized with one virus-secreting cell placed in the middle of a 2-dimensional sheet of hexagonally-tiled epithelial cells (288,212 total cells).  The sheet measures 5$mm$ per side, which is large enough to contain the infection (Fig.~\ref{fig:cycells}).  The molecular grid (Text S1 1.3) is initialized with a resolution of 20$\mu m$.  Each time step of the model represents 6 seconds.

@@

\subsection{Modeling decisions}

@@

The model simulates a chemokine gradient surrounding the infected focus (Fig.~\ref{fig:cycells}), based on the calculated per-infected cell secretion rate (Table \ref{tab:strains}) and known chemical parameters for a 10 kDa protein (Table \ref{tab:parameters}).  T cell sensitivity depends on receptor density \citep{Desmetz2006} and this was assumed to be constant.

@@

% we hypothesize that it renders IP-10 ineffectual.  We hypothesize that this leads to the elevated RANTES secretion rates measured in aH5N1 compared to the other two strains (Table \ref{tab:strains}).  Because of this behavior IP-10 was not included in the aH5N1 model runs.

@@

\subsection{One-factor-at-a-time sensitivity analysis}

@@

The model parameters were chosen from literature when available and estimated within plausible ranges otherwise (Tables \ref{tab:strains} and \ref{tab:parameters}).  One-factor-at-a-time (OFAT) analysis was performed by varying Individual parameters over ranges of plausible (and sometimes even implausible) values while the rest of the parameter set was held constant.  Each parameter was varied over all three influenza strains, creating three sets of sensitivity plots (Figures~\ref{fig:asensitivity}-\ref{fig:psensitivity}).

We then categorized the model parameters into one of three qualitative groups: parameters that do not affect the model's qualitative behavior (stable parameters), parameters that affect the model's quantitative peak infection size but do not affect final clearance, (peak change parameters), and those parameters that affect the final clearance of the infection (sensitive parameters) (Table \ref{tab:sensitivity}).  Parameters are presented along with their range using the following format:  \\

@@

Stable parameters do not affect the outcome of the infection unless they are adjusted to values outside the realm of possibility.  Specifically, the plausible parameter values in this group are within a range that does not significantly affect the asymptotic behavior of the model.  We further split this group into two categories in Table \ref{tab:sensitivity}, \textit{stable} parameters and \textit{bounded stable} parameters.  \textit{Bounded stable} parameters are only stable within a specific range of values, while \textit{stable} parameters seem to be stable over all the values that were tested.

@@

Compartmental modeling design suggests the removal of such stable parameters.  Due to the nature of the ABM, inclusion of these parameters is mechanistically necessary.  Thus, we include and parameterize them using the methods described in the main paper and Table \ref{tab:parameters}.

@@

\textbf{T Cell Age in Blood:} [ \textit{1h |  4d | 400d} ] - The expected age determines the decay rate of T cells in blood.

@@

\textbf{T Cell Age at FOI:} [ \textit{12m | 2h | 20h} ] - The expected age determines the decay rate of T cells at the focus of infection (FOI).

@@

An explicit sensitivity analysis was not directly performed.  The range of the parameters are taken from the respective secretion rates of the aH5N1, sH1N1, and pH1N1 influenzas (Table \ref{tab:strains}).

@@

\subsection{Partial rank correlation coefficient sensitivity analysis}

We performed a partial rank correlation coefficient (PRCC) sensitivity analysis to complement our one factor at a time (OFAT) parameter analysis.  Due to computational limitations, we computed PRCC values only for the sH1N1 strain, confirming that the results of the PRCC are consistent with the conclusions of the original sensitivity analysis and that viral parameters dominate the model output.

\subsubsection{PRCC Design and Implementation}

PRCC analysis was performed as described in \cite{Marino2008}.  Samples were generated using Latin hypercube sampling (LHS) to generate 333 sample points for 17 parameters (those from Table \ref{tab:sensitivity}, plus the sH1N1 viral secretion rate).  Each parameter was assigned a confidence distribution (Table \ref{tab:lhs}) and sampled using its inverse cumulative distribution function to generate points with density proportional to the underlying probability distribution function.  Each sample point was evaluated three times using unique random seeds for a total of 999 samples.  One sample point resulted in undefined model behavior and was removed, leaving a total of 996 samples.

Output is defined as the number of infected cells remaining for each time point.  Each model run generates 144,001 time points (10 day model with 6 second time steps).  PRCC analysis was performed over every 100th time point, resulting in a time series of 1441 Spearman rank correlation coefficients ($\rho$) and significance values ($p$) for each parameter (Figs. \ref{fig:prcc} - \ref{fig:vsec}).

\subsubsection{PRCC Results}

The  analysis confirmed the conclusion of original sensitivity analysis that viral parameters dominate model output.  Of the 17 parameters tested with PRCC analysis, 7 show significance of $p < 0.01$ over the time period where the parameter was relevant to model dynamics (Figs. S6 - S7).  All five of the viral parameters (viral response to IgM, viral infectivity, viral decay rate, viral diffusion rate, and viral secretion rate) are significant.  Infected cell expression time and the T cell age at the focus of infection are also significant at $p < 0.01$ (but have $\rho$ values near 0).  Viral infectivity, viral decay rate, viral diffusion rate, and viral secretion rate are the only parameters that have maximum absolute vales of $\rho$ above 0.5.  Other parameters showed significant correlations (with low values of $\rho$), but only over time periods where they were not active.

Viral infectivity, viral diffusion rate, and viral secretion rate, values of $\rho$ cross or approach zero over time.  This merely reflects that the modeled target cells have died (Fig.~\ref{fig:exhaustive}) and so the decline in $\rho$ is irrelevant.

%This effect is due to the spatial constraints built into our agent based model.  The  model environment was designed to be large enough to contain an infection defined by the default model parameter values.  Extreme values of the viral parameters cause rapid infection growth to the point that the modeled system is exhausted (Fig. \ref{fig:exhaustive}).  In these cases the number of infected cells tends towards zero at later time points as every modeled target cell has died before the model is complete.  Thus, the calculated values of $\rho$ for these given parameters should be trusted only at earlier time points.

@@

\begin{figure}[!ht]
\begin{center}
 \includegraphics[width=\textwidth]{FigureS1}
 \end{center}
\caption{{\bf Preliminary Model 1 fits to data.}  Preliminary Model 1 (Eq. 1) was fit to experimental data (Table~\ref{tab:data}) using a genetic algorithm as described in the main paper.  sH1N1 IP-10 secretion exceeded measurement accuracy above 8500 pg/mL and these three values (empty red triangles) were not included in the model fitting.  RMSE values of the fits can be seen in Table~\ref{tab:rmse}.  IP-10 data and model fits shown in red (triangles), RANTES data and model fits shown in yellow (squares).} 
 \label{fig:fits}
\end{figure}

\begin{figure}[!ht]
\begin{center}
 \includegraphics[width=\textwidth]{Figure_S2}
 \end{center}
\caption{{\bf Varying T cell sensitivity to chemokine.}  H5N1 model results use RANTES  only, and sH1N1 and pH1N1 use both IP-10 and RANTES. Total number of incubating, secreting and apoptotic cells are plotted for each infection.  The sensitivity value specifies the minimum level of chemokine concentration required for T cells to detect it. } 
 \label{fig:sensitivity}
\end{figure}

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=\textwidth]{Figure_S3}
	\caption{\textbf{Effects of different chemokine combinations.}  A) aH5N1 does not stimulate an IP-10 response.  B-C) sH1N1 and pH1N1 show no significant difference between IP-10 alone versus IP-10 and RANTES combined.}
	\label{fig:chemokine}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=9in]{Figure_S4}
	\caption{\textbf{aH5N1 sensitivity analysis.} The sensitivity analysis results for 16 parameters applied to the aH5N1 model simulation.  In each subgraph, the thicker purple line shows the results of the model using the default parameter value.}
	\label{fig:asensitivity}
\end{center}
\end{figure}

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=9in]{Figure_S5}
	\caption{\textbf{sH1N1 sensitivity analysis.} The sensitivity analysis results for 16 parameters applied to the sH1N1 model simulation.  In each subgraph, the thicker purple line shows the results of the model using the default parameter value.  The divergence between Apoptosis Time values of 90m and 120m (upper-left plot) highlights an important threshold of infection control.}
	\label{fig:ssensitivity}
\end{center}
\end{figure}

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=9in]{Figure_S6}
	\caption{\textbf{pH1N1 sensitivity analysis.} The sensitivity analysis results for 16 parameters applied to the pH1N1 model simulation.  In each subgraph, the thicker purple line shows the results of the model using the default parameter value.}
	\label{fig:psensitivity}
\end{center}
\end{figure}

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=9in]{Figure_S7}
	\caption{\textbf{PRCC sensitivity analysis.} PRCC analysis generated Spearman correlation coefficients ($\rho$) and significance values ($p$) over 1441 time points.  Plots show values of $\rho$ vs. time for each parameter tested.  Dark blue indicates significance levels at $p < 0.01$.  Gold indicates significance of $p < 0.05$.  Light gray indicates $p \geq 0.05$.  Six parameters show significant effect over the time period where they were relevant to model dynamics: infected cell expression time, viral response to IgM, viral infectivity, T cell expected lifetime at the FOI, viral decay rate, and the viral diffusion rate.}
	\label{fig:prcc}
\end{center}
\end{figure}

@@

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=\textwidth]{Figure_S8}
	\caption{\textbf{PRCC sensitivity analysis for the viral secretion rate.} Similar to Figure S6, PRCC analysis generated Spearman correlation coefficients ($\rho$) and significance values ($p$) over 1441 time points.  Plot shows the values of $\rho$ vs. time for the viral secretion rate.  Dark blue indicates significance levels at $p < 0.01$.  Gold indicates significance of $p < 0.05$.  Light gray indicates $p \geq 0.05$.  Viral secretion shows significant effect on model output.  The decline of $\rho$ over time reflects the early death of every modeled target cell (Figure~\ref{fig:exhaustive})).}
	\label{fig:vsec}
\end{center}
\end{figure}

\begin{figure}[ht!]
\begin{center}
	\includegraphics[width=\textwidth]{Figure_S9}
	\caption{\textbf{Exhaustive Growth:} Certain parameter sets generated by Latin hypercube sampling caused rapid uncontrolled infection growth (Samples 1-3).  This resulted in the early death of every modeled target cell (saturation at day 1) and the inevitable decline of the infection size to zero.  A larger, unbounded model environment would show unconstrained growth.  This effect explains the decline of the $\rho$ value of several viral parameters to or past zero over time in Figures \ref{fig:prcc} and \ref{fig:vsec}.}
	\label{fig:exhaustive}
\end{center}
\end{figure}

%\begin{figure}[ht!]
%\begin{center}
%	\includegraphics[width=\textwidth]{Figure_SX}
%	\caption{\textbf{pH1N1 sensitivity analysis.} Model results and fits of the viral effectiveness rate, $E$.}
%	\label{fig:effectiveness}
%\end{center}
%\end{figure}

@@

\begin{table}[!ht]
\begin{center}
\begin{tabular}{|  l  l  l  |}
  \hline
  Population & Description & Initial Value\\
  \hline
  $T$ & Healthy target epithelial cells & 1,000,000 \\
  $I_1$ & Virus-incubating cells & 0 \\
  $I_2$ & Virus-secreting cells & 0 \\
  $V$ & Free virus particles & 10,000 \\
  $F$ & Interferon quantity & 0 \\
  $C$ & Chemokine quantity (ng/mL) & 0 \\
  \hline
  \hline
  Parameter & Description & Value \\
  \hline
  $\beta$ & Viral infection rates &  \\
    & \hspace{2em} Avian & 5.3e-7 (PFU$\cdot$h)$^{-1}$ \\
    & \hspace{2em} Seasonal & 6.1e-7 (PFU$\cdot$h)$^{-1}$ \\
    & \hspace{2em} Pandemic &  2.7e-6 (PFU$\cdot$h)$^{-1}$\\
  $p$ & Viral production rates & \\
    & \hspace{2em} Avian & 0.20 (PFU/h) \\
    & \hspace{2em} Seasonal & 1.4 (PFU/h) \\
    & \hspace{2em} Pandemic & 18.3 (PFU/h) \\
  $e$ & Interferon strengths & \\
    & \hspace{2em} Avian & 1.0e-8 \\
    & \hspace{2em} Seasonal & 1.6e-6 \\
    & \hspace{2em} Pandemic & 3.4e-3 \\
  $\tau_2$ & Interferon production delays &  \\
    & \hspace{2em} Avian & 21.5 (h) \\
    & \hspace{2em} Seasonal & 23.6 (h) \\
    & \hspace{2em} Pandemic & 21.0 (h) \\
  $\delta$ & Virus-secreting cell decay rate & 0.6 (h$^{-1}$) \\
  $d$ & Chemokine decay rate & 1.386 (h$^{-1}$) \\
  $\tau_1$ & Viral incubation time & 10 (h) \\
  $\tau_3$ & Chemokine production delays & \\
    & \hspace{2em} IP-10 & 8 (h)\\
    & \hspace{2em} RANTES & 16 (h)\\
  \hline
\end{tabular}
\caption{\textbf{Parameters for Preliminary Model 1 (Equation 1)}.  All parameters and populations are taken from \citep{Mitchell2011} except for $C$, $d$, and $\tau_3$.  The value chosen for $d$ corresponds to a 30 minute half-life and values for $\tau_3$ were set from the observed chemokine data (Fig.~\ref{fig:data}).  Interferon ($F$) is an abstracted unitless quantity and thus $e$ is a unitless multiplier.}
\label{tab:dde}
\end{center}
\end{table}

\begin{table}[!ht]
\begin{center}
\begin{tabular}{| c | c | c |}
  \hline
  Strain & Chemokine & RMSE \\
  \hline
  \multirow{2}{*}{aH5N1} & RANTES & 166 \\
  & IP-10 & 7,950 \\
  \hline
  \multirow{2}{*}{sH1N1} & RANTES & 90 \\
  & IP-10 & 2,732 \\
  \hline
  \multirow{2}{*}{pH1N1} & RANTES & 2,671  \\
  & IP-10 & 64,342 \\
  \hline
\end{tabular}
\caption{\textbf{Root mean squared error of Preliminary Model 1 fits.}  Preliminary Model 1 (Eq. 1) was fit to experimental data (Table~\ref{tab:data}) using a genetic algorithm as described in the main paper.  The resulting model fits can be seen in Figure~\ref{fig:fits}.}
\label{tab:rmse}
\end{center}
\end{table}

\begin{table}[!ht]
\begin{center}
\begin{tabular}{ | c | c | c |}
  \hline                        
  Paramter & Value & Source \\
  \hline
  T(0) & $1e6$ & \citep{Mitchell2011} \\
  V(0) &  $1e4$ & \citep{Mitchell2011} \\
  $r_{cell}$ &  $5 \mu m$ & \citep{Miao2010a} \\
  $t_{rc}$ & $6s$ & \citep{Peters1983} \\
  $k_e$ & $6.4e-5/cell/day$ & \citep{Miao2010a} \\
  $\beta$ & $4.8e-7/cell/PFU$ & \citep{Mitchell2011} \\
  $p$ & $0.18 PFU/h$ & \citep{Mitchell2011} \\
  $\delta$ & $16.7 hours$ & \citep{Mitchell2011} \\
  $a$ & $67.7$ & ABM fits \\
  $b$ & $433.5$ & ABM fits \\
  \hline  
\end{tabular}
\begin{tabular}{| c | c |}
  \hline
  I & C \\
  \hline
  4 & 329 \\
  58 & 997 \\
  816 & 2550 \\
  3,264 & 4,410 \\
  10,050 & 8,069 \\
  \hline
\end{tabular}
\caption{\textbf{Parameters for Preliminary Model 2 (Equation 2).}  Values are chosen to match the equations borrowed from both \citep{Mitchell2011} and \citep{Miao2010a}.  Values \textit{a} and \textit{b} were fit to multiple runs of a simplified CyCells ABM using the linear model $C = a\sqrt{I}+b$ with a resulting $R^2 = 0.997$, $P < 0.01$.}
\label{tab:supplement}
\end{center}
\end{table}

\begin{table}[!ht]
\centering
\begin{tabular}{| c | l | c | l |}
  \hline                        
  Category & Parameter & Units & Distribution \\
  \hline
  \multirow{3}{*}{Chemokine} & Chemokine Decay Rate & \textit{Hz} &  LN(-3.4, 1.0) \\
  & Chemokine Diffusion Rate & \textit{$\mu$m$^2$/s} & LN(-0.5, 1.0) \\
  & Chemokine Secretion Rate & $(pg/s\cdot cell)$ & LN(-3.7, 0.1) \\
  \hline
  \multirow{6}{*}{T Cell} & Circulation Time & \textit{seconds} & LN(-.75, 0.5)\\
  & T Cell Kill Rate & \textit{min} & LN(1.0, 0.5) \\
  & T Cell Speed & \textit{$\mu$m/min} & LN(-.75, 1), \\
  & T Cell Age in Blood & \textit{days} & LN(0.6, 1.0) \\
  & T Cell Age at FOI& \textit{min} & LN(2.1, 0.5) \\
  & T Cell Production Rate & \textit{cells/h} & LN(3.1, 0.15) \\
  \hline
  \multirow{3}{*}{Delay} & Apoptosis Time & \textit{hours} & U(0.0, 2.0) \\
  & Expression Time & \textit{min} & LN(3.0, 0.15) \\
  & Incubation Time & \textit{hours} & LN(-1.0, 0.15) \\
  \hline 
  \multirow{4}{*}{Virus} & Viral Response to IgM & | & LN(1.0, 0.5) \\
  & Infectivity & \textit{min/virion} & LN(2.1, 0.5) \\
  & Viral Decay Rate &  \textit{day$^{-1}$} & LN(0, 1.0) \\
  & Viral Diffusion Rate & \textit{$\mu$m$^2$/s} & LN(-1.5, 1.0) \\
  & Viral Secretion & $(PFU/s\cdot cell)$ & LN(-3.4, 0.3) \\
  \hline  
\end{tabular}
\caption{\textbf{Latin hypercube sampling distributions.}  LHS was used to generate sample points containing each of the 16 parameters from Table \ref{tab:sensitivity}, plus the sH1N1 viral secretion rate.  Parameters were sampled over biologically plausible ranges listed here for use in PRCC analysis.  LN signifies a log-normal distribution.  Values were sampled from a regular normal distribution with the listed parameters and then transformed to linear space from $log_{10}$ space.  U signifies a uniform distribution.}
\label{tab:lhs}
\end{table}

### 2016-01-28 paper/cover_letter.tex
In process of drafting Response 3

\begin{flushleft}Dear Editors, \end{flushleft}

We submit this manuscript entitled "Spatially explicit model of the efficiency of the lymphocyte diaspora in influenza-infected lung reveals thresholds on chemokine directed migration" for consideration of publication in PLOS ONE.  Our paper implements an agent-based model of bronchial influenza infection and the subsequent T cell response using new empirical values of chemokine secretion.  The paper shows that chemokine levels have a limited yet important role in directing T cell migration and that the spatial effects of chemokine diffusion can actually hinder T cell search in rapidly spreading viruses.  We feel this work will be of interest to investigators modeling the immune response for the ultimate purpose of improving vaccine design.
 \\

This work has not been published elsewhere, and is complementary to a recent publication from our group (H Mitchell, D Levin, et al. J Virol 2011;85(2):1125-1135). \\

In review we have noticed that our bibliography does not use annotated journal names, which seem to be the standard for PLOS ONE.  We have {\em not} changed our bibliography as of yet as it was generated automatically using the Mendeley Desktop suite.  If this is an issue, we would be happy to return and reconstruct the references manually. \\

% This work is a resubmission of a rejected PLOS CB manuscript (PCOMPBIOL-D-12-01516R1).

All authors have read and approved the revised manuscript.

\begin{flushleft}Respectfully yours,\end{flushleft}
Drew Levin \\
Stephanie Forrest \\
Soumya Banerjee \\
Candice Clay \\
Melanie Moses \\
Frederick Koster

>>>>

\pagenumbering{gobble}

\begin{flushleft}Dear Dr. Denise Kirschner \\
Co-Editor-in-Chief \\
Journal of Theoretical Biology, \end{flushleft}

%We submit this manuscript entitled "A spatial model of the efficiency of T cell search in the influenza-infected lung" for consideration of publication in Royal Society Interface.  Our paper implements an agent-based model of bronchial influenza infection and the subsequent T cell response using new empirical values of chemokine secretion.  The paper shows that chemokine levels have a limited yet important role in directing T cell migration and that the spatial effects of chemokine diffusion can actually hinder T cell search in rapidly spreading viruses.  Our sensitivity analysis reveals the individual contributions of the model parameters on the course of the infection.  Finally, we calculate a maximum threshold for T cell search time, beyond which leads to uncontrolled infection.  We feel this work will be of interest to investigators modeling the immune response for the ultimate purpose of improving vaccine design.

We are pleased to submit the revised manuscript, ``A spatial model of the efficiency of T cell search in the influenza-infected lung" for JTB's consideration. This is a resubmission of manuscript number JTB-D-15-00501, received by us on August 12, 2015 with an invitation for resubmission within 60 days. \\

The reviewer comments were extremely helpful, and we have addressed each of their suggestions in the attached Response to Reviewers.   To summarize the more important changes, we performed the recommended PRCC sensitivity analysis for one viral strain. Because it confirmed the results of our original sensitivity analysis, we conducted this computationally intensive analysis for only this strain.  Beyond the PRCC runs, we doubled the number of model runs using our standard parameters.  This resulted in a very small decrease in the variance of the results and did not affect the mean.  This gives us additional confidence in the validity of the reported results.  Next, we reorganized the Methods and Model section to make the rationale for our parameter choices more clear.  We note that the PRCC sensitivity analysis confirms that many of these parameters do not have a strong effect on the model output.  Finally, we corrected the typographical errors pointed out by the referees.   \\

We feel that the paper is much improved with these changes, and we appreciate your patience during the revision process.

\begin{flushleft}Sincerely, \end{flushleft}

\noindent Drew Levin \\
University of New Mexico Department of Computer Science \\
MSC01 1130 \\
1 University of New Mexico \\
Albuquerque, NM 87131 \\
drew@cs.unm.edu
